Hilbert space methods to approximate Gaussian processes using Stan

### 2019-03-21 Paper/manuscript/manuscript_05.tex
improve section 3 and figures

% This graph basically provides the recommended values for these factors in order to get a close approximation. This graph also allows us to diagnose an actual fit by checking whether the value of the estimated lengthscale does not keep to those that the Figure brings as suitable ones.

%We compare regular GP and HSGP fitted for different dataset with different "wigglyness".
%The comparison is made through comparing their estimated lengthscales.

@@

The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP. The approximation also depends on the boundary $L$ (see Equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})), which will affect its accuracy especially near the boundaries. As we wil see later on, $L$ will also influence the number of basis functions required in the approximation. In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi$. Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \mathbb{R}$. We now define $L$ as
%
\begin{equation}\label{eq:boundary}
L=c \cdot S,
\end{equation}

@@

We start with an illustration on how the number of basis functions $m$ and boundary factor $c$ influences the accuracy of the HSGP approximations, separately. For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ using input values from the zero-mean input domain with half-range $S=1$. Several HSGP models with varying $m$ and $L$ are fitted to this data. In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model. 
Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictions of the estimated function and on the covariance function itself. For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the non-linearity of the estimated function: Fewer basis functions imply smoother covariance functions (Paul: when would you consider a covariance function as smooth? Also, I think the main point here is that at long distances the implied covariance function becomes negative which can't be in an exact GP model) and consequently more linear functional forms. The higher the "wigglyness" of the function to be estimated, the more basis functions will be required (Paul: do we see this in Figure \ref{fig1_Post_J} already?). For $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.43, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
\caption{(left) Posterior distributions for different number of basis functions $m$. (right) Covariance function for different number of basis functions $m$.}
  \label{fig1_Post_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.43]{fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.43]{fig2_Cov_L.pdf}}
\caption{(left) Posterior means for different different values of the box size $L$. (right) Covariance functions for different values of the box size $L$.}
  \label{fig2_Post_L}
\end{figure}

Next, we will focus on analyzing the interaction effects between these $m$ and $c$ on the performance of the approximation. The lengthscale and marginal variance will no longer be fixed but rather estimated in both regular GP and HSGP models.  
%The right-side plots in Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} are computed using the lengthscale and magnitud values of the true GP prior. Now, we focus on the estimated lengthscale and magnitud parameters after fitting the model, either of the regular GP model fit and approximate GP model fit. 
Figure \ref{fig4_Post_&_Cov_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$. Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the approximate GP model, computed against the regular GP model. Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models. Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$. Further, the choice of $m = 10$ and $c = 1.5$ could still be an accurate enough choice. We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}. From these results, some general conclusions may be drawn:

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part1.pdf}}
\subfigure{\includegraphics[scale=0.6]{fig4_Post_&_Cov_part2.pdf}}
\caption{Posterior distributions and covariance functions of the regular GP model and the proposed HSGP model as a function of the number of basis functions $m$ and the boundary factor $c$.} 
  \label{fig4_Post_&_Cov_part1}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
\caption{Standardized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model. (left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$. (right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig5_MSE_vs_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4]{fig6_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig6_magnitud_vs_J.pdf}}
\caption{Estimated lengthscale and marginal variance parameters of the proposed HSGP models and the regular GP model. (left) Estimated lengthscales vs the number of basis functions $m$ and for different values of the boundary factor $c$. (right) Estimated marginal variance vs the number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig6_lscale_vs_J}
\end{figure}

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figure \ref{fig3_lscale_vs_J_vs_c_part1} depicts how these three factors interact with each other in relation to the performance of the HSGP approximation (Paul: Did you say how you choose $\varepsilon$?). On the X-axis of this plot is placed the lengthscale of the process normalized by the half-range of the data $\ell/S$. The countour lines gather the boundary factor $c$. And the Y-axis represents the number of basis functions (Paul: This needs to be put in the figure caption not in the text). More precisely, for a given GP function with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig3_lscale_vs_J_vs_c_part1} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}). Alternatively, this figure could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$. Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$. We obtain the following main conclusions:

\begin{itemize}
\item As $\ell$ increases, $c$ and $m$ required for a close enough approximation decrease.
\item There is a minimum $c$ under which a close approximation is never going to be achieved (The contour lines have an end in function of the lengthscale; Paul: what do you mean?).
\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve a close approximation.
\end{itemize}

For users of HSGP models, Figure \ref{fig3_lscale_vs_J_vs_c_part1} also severs as a powerful diagnostic tool in determining if the obtained accuracy is acceptable.  Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ is exceeds the corresponding line provided in Figure \ref{fig3_lscale_vs_J_vs_c_part1}. If $\hat{\ell}$ exceeds the line, the approximation should be close enough. If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased (Paul: decreasing $c$ seems dangerous to me if predictions for out of sample data are desired. Shell we mention this?). We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can fit an HSGP model, check the approximation accuracy, and, if not accurate enough, repeat the process while replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
\caption{Relation between the minimum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
  \label{fig3_lscale_vs_J_vs_c_part1}
\end{figure}

If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig3_lscale_vs_J_vs_c_part1}.

@@

Another way to look at the approximation accuracy of HSGPs is to compare the lengthscale estimates
obtained from the regular GP and HSGP models. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well.

For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3?) and varying lengthscale values. Different values of the number of basis functions $m$ are used when estimating the HSGP models while the boundary factor $c$ is fixed to some reasonable value. Figure \ref{fig7_varing_lscale} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
\caption{Posterior means of the regular GP and approximate GP models fitted over different datasets with different lengthscales.}
  \label{fig7_varing_lscale}
\end{figure}

The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}. In the X-axes of the plots, the true lengthscales of the data-generative functions for every dataset scenario are represented. In the Y-axes, the estimated lengthscales for the regular GP and approximate GP models are represented (Paul: This needs to go into the Figure captions). The dashed black line represents the minimum lengthscales that can be closely approximate given the number of basis functions and the boundary factor, according to Figure \ref{fig3_lscale_vs_J_vs_c_part1} (This also needs to go into the Figure caption, but may be worth keeping in the text as well). Further, Figure \ref{fig9_MSE_varing_lscale} shows the SRMSE (Paul: you did not define that anywhere before) of the HSGP against the regular GP model as a function of the lengthscale and number of basis functions.

We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales. Very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high SRMSE. On the other side, big changes on long lengthscales, which might clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small SRMSE.

(Paul: I don't get the last paragraph. Can you rewrite this a little bit yourself and also merge it with the former paragraph? I am happy to look at it again afterwards.)

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
\caption{True lengthscales versus estimated lengthscales of the different realizations of datasets using different number of basis functions. }
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
\caption{SMSE of the approximate GP model for the different datasets with different wiggly effects and number of basis functions.}
  \label{fig9_MSE_varing_lscale}
\end{figure}

>>>>

@@

The finite number $m$ of basis functions in the approximation needed to satisfy Equation (\ref{diff_covs}) depends on the non-linearity of the function to be learned, that is on its lengthscale $\ell$, which constitutes a hyperparameter of the GP. The approximation also depends on the boundary $L$ (see Equations (\ref{bf&lambda}) and (\ref{bf&lambda_multi})), which will affect its accuracy especially near the boundaries. As we will see later on, $L$ will also influence the number of basis functions required in the approximation. In the present paper, we will set $L$ an extension of the desired covariate input domain $\Psi$. Without loss of generality, we can assume $\Psi$ to be symmetric around zero, that is $\Psi=[-S,S] \subset \mathbb{R}$. We now define $L$ as
%
\begin{equation}\label{eq:boundary}
L=c \cdot S,
\end{equation}

@@

We start with an illustration on how the number of basis functions $m$ and boundary factor $c$ influences the accuracy of the HSGP approximations, separately. For this purpose, a set of noisy observations are drawn from an exact GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$ using input values from the zero-mean input domain with half-range $S=1$. Several HSGP models with varying $m$ and $L$ are fitted to this data. In this example, the lengthscale and marginal variance parameters used in the HSGPs are fixed to the true values of the data-generating model. 
Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of $m$ and $c$, respectively, on the posterior predictions of the estimated function and on the covariance function itself. For $c$ fixed to a large enough value, Figure \ref{fig1_Post_J} shows clearly how $m$ affects the accuracy on the approximation and the non-linearity of the estimated function, in the sense that fewer basis functions implies inaccurate and smoother effects covariance functions (covariance functions with larger lengthscales) and consequently more linear functional forms, and conversely. 
%:Fewer basis functions imply smoother covariance functions (Paul: when would you consider a covariance function as smooth? Also, I think the main point here is that at long distances the implied covariance function becomes negative which can't be in an exact GP model.) and consequently more linear functional forms. 
Then, the higher the "wigglyness" of the function to be estimated, the more basis functions will be required. 
%(Paul: do we see this in Figure \ref{fig1_Post_J} already?). 
On the other hand, for $m$ fixed to a large enough value, Figure \ref{fig2_Post_L} shows that $c$ mainly affects the approximation near the boundaries as well as covariances at long distances.

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.47, trim = 0mm 0mm 0mm 10mm, clip]{fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.47, trim = 25mm 0mm 0mm 10mm, clip]{fig1_Cov_J.pdf}}
\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different number of basis functions $m$. (right) Covariance functions of both regular GP model and HSGP model for different number of basis functions $m$.}
  \label{fig1_Post_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.47, trim = 0mm 0mm 0mm 10mm, clip]{fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.47, trim = 25mm 0mm 0mm 10mm, clip]{fig2_Cov_L.pdf}}
\caption{(left) Posterior predictive mean functions of both regular GP model and HSGP model for different boundary factors $c$. (right) Covariance functions of both regular GP model and HSGP model for different boundary factors $c$.}
  \label{fig2_Post_L}
\end{figure}

Next, we will focus on analyzing the interaction effects between these $m$ and $c$ on the performance of the approximation. The lengthscale and marginal variance will no longer be fixed but rather estimated in both regular GP and HSGP models. Figure \ref{fig4_Post_&_Cov_part1} shows the functional posterior predictions and the covariance function obtained after fitting the data, for varying $m$ and $c$. Figure \ref{fig5_MSE_vs_J} shows the root mean square error (RMSE) of the HSGP models, computed against the regular GP model. Figure \ref{fig6_lscale_vs_J} shows the estimated lengthscale and marginal variance for the regular GP model and the HSGP models. Looking at the RMSEs in Figure \ref{fig5_MSE_vs_J}, we can conclude that the optimal choice in terms of precision and computations would be $m = 15$ basis functions and a boundary factor between $c = 1.5$ and $c = 2.5$. Further, the choice of $m = 10$ and $c = 1.5$ could still be an accurate enough choice. We may also come to the same conclusion by looking at the posterior predictions and covariance function plots in Figure \ref{fig4_Post_&_Cov_part1}. From these results, some general conclusions may be drawn:

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.41, trim = 0mm 3mm 0mm 0mm, clip]{fig4_Post_&_Cov_part1.pdf}}
\subfigure{\includegraphics[scale=0.41, trim = 0mm 3mm 0mm 3mm, clip]{fig4_Post_&_Cov_part2.pdf}}
\subfigure{\includegraphics[scale=0.41, trim = 0mm 0mm 0mm 3mm, clip]{fig4_Post_&_Cov_part3.pdf}}
\caption{Posterior predictive mean functions and covariance functions of both regular GP model and HSGP model for different number of basis functions $m$ and values of the boundary factor $c$.} 
  \label{fig4_Post_&_Cov_part1}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.44]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.44]{fig5_MSE_vs_c.pdf}}
\caption{Root mean square error (RMSE) of the proposed HSGP models computed against the regular GP model. (left) RMSE versus the number of basis functions $m$ and for different values of the boundary factor $c$. (right) RMSE versus the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig5_MSE_vs_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.48, trim = 0mm 0mm 5mm 0mm, clip]{fig6_lscale_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.48, trim = 31mm 0mm 0mm 0mm, clip]{fig6_magnitud_vs_J.pdf}}
\caption{Estimated lengthscale (left) and marginal variance (right) parameters of both regular GP and HSGP models, plotted versus the number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig6_lscale_vs_J}
\end{figure}

Additionally, there is a clear relation of the number of basis functions $m$ and the boundary factor $c$ with the lengthscale $\ell$ of the approximated function. Figure \ref{fig3_lscale_vs_J_vs_c} depicts how these three factors interact with each other in relation to 
% the performance of the HSGP approximation.
a close approximation of the HSGP model, in the case of a GP with square exponential covariance function and a single input dimension. (Gabi: Could it be good to build and include the same Figure for a Matern($\frac{3}{2}$)?). More precisely, for a given GP model (with a square exponential covariance function) with lengthscale $\ell$ and given a boundary factor $c$, Figure \ref{fig3_lscale_vs_J_vs_c} shows the minimum $m$ required to achieve a close approximation in terms of satisfying Equation (\ref{diff_covs}). We have considered an approximation to be a close enough when the difference between densities of the approximate covariance function and the exact covariance function, $\varepsilon$ in Equation (\ref{diff_covs}), is below 1$\%$ of the density of the exact covariance function,
%
\begin{eqnarray*}
 \frac{\varepsilon}{\int k(\tau) \, d\tau} < 0.01.
\end{eqnarray*}

\noindent Alternatively, this figure could be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$. Of course, we may also read it as providing the minimum $\ell$ that can be closely approximated given $m$ and $c$. We obtain the following main conclusions:

\begin{itemize}
\item As $\ell$ increases, $c$ and $m$ required for a close enough approximation decrease.
\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve a close approximation.
\item For a given $\ell$ there exist a minimum $c$ under which a close approximation is never going to be achieved regardless of $m$. This fact can be appreciated in the Figure as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$).
\end{itemize}

As stated above, Figure \ref{fig3_lscale_vs_J_vs_c} provides the minimum lengthscale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the lenghscale $\ell$ controls the "wigglyness" of
the functional relationship, it strongly influences the difficulty of obtaining accurate
inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
we can expect the HSGP approximation to be accurate as well. Thus, having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum lengthscale provided in Figure \ref{fig3_lscale_vs_J_vs_c}. If $\hat{\ell}$ exceeds this recommended minimum lengthscale, the approximation should be close enough. If, however, it does not exceed it, the approximation may be inaccurate and $m$ should be increased or $c$ decreased (Paul: decreasing $c$ seems dangerous to me if predictions for out of sample data are desired. Shell we mention this? Gabi: We assume that predicting points are already within $S$. Then, a valid $c$ is more related to the lengthscale than to the predicting points. In fact, valid $c$s are restricted in function of the lengthscale: A given lengthscale has a minimum $c$ under which a close approximation will never be achieved regarless $m$ (We stated this in line 502 and also in line 462). For example, a given $c$ can not work for certain lengthscale but for a larger one). We may also use this diagnostic in a iterative procedure.
Starting from some initial guess of $\ell$, we can choose initial values for $m$ and $c$ and fit an HSGP model, then check the approximation accuracy, and, if not accurate enough because the estimated $\hat{\ell}$ is below the minimum lengthscale provided by Figure \ref{fig3_lscale_vs_J_vs_c}, repeat the process while increasing $m$ or decreasing $c$. Note that, as commented before, $c$ can not be decreased as much as desired because it is restricted to the lengthscale.
%replacing $\ell$ by the newly obtained estimate $\hat{\ell}$.

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.44]{fig3_lscale_vs_J_vs_c.pdf}}
\subfigure{\includegraphics[scale=0.44, trim = 0mm 0mm 0mm 0mm, clip]{fig3_lscale_vs_J_vs_c_zoomin.pdf}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ ($c = \frac{L}{S}$), and the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$). The right-side plot is a zoom in of the left-side plot.}
  \label{fig3_lscale_vs_J_vs_c}
\end{figure}

If we look back to the conclusions drawn from Figures \ref{fig5_MSE_vs_J} and \ref{fig6_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were enough to closely approximate a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure \ref{fig3_lscale_vs_J_vs_c}.

(Gabi: The following paragraph is just an idea to extend the discussion to the multidimensional case)

Figure \ref{fig3_lscale_vs_J_vs_c} was build for a GP with a unidimensional covariance function, and results in a surface depending of three variables, $m$, $c$ and $\ell$. An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending of four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which can not be directly represented. However, as an approximation we can use the unidimensional GP conclusions in Figure \ref{fig3_lscale_vs_J_vs_c} to analyze individually the different dimensions of a multidimensional GP model.

@@

In this example, we make a comparison of the lengthscale estimates
obtained from the regular GP and HSGP models. We also have a look at those recommended minimum lengthscales provided by Figure \ref{fig3_lscale_vs_J_vs_c}.

%Another way to look at the approximation accuracy of HSGPs is to compare the lengthscale estimates
%obtained from the regular GP and HSGP models. As the lenghscale $\ell$ controls the "wigglyness" of
%the functional relationship, it strongly influences the difficulty of obtaining accurate
%inference about the function from the data. Basically, if the lengthscale estimate is accurate, 
%we can expect the HSGP approximation to be accurate as well.

For this analysis, we will use various datasets consisting of noisy draws from a GP prior model with a squared exponential covariance function (Paul: Did we talk about the applied kernels previously in Section 3? Gabi: I've mentioned it in line 491) and varying lengthscale values. Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case.

Figure \ref{fig7_varing_lscale} shows the posterior predictions of the regular GP and HSGP models fitted to those datasets. The Lengthscale estimates as obtained by regular GP and HSGP models are depicted in Figure \ref{fig8_Tlscale_vs_Elscale}. Further, Figure \ref{fig9_MSE_varing_lscale} shows the root mean square error (RMSE) of the HSGP models against the regular GP models as a function of the lengthscale and number of basis functions $m$.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig7_varing_lscale.pdf}
\caption{Posterior predictive means of both regular GP and HSGP models, fitted over different datasets simulated from square exponential GP models with different characteristic lengthscales ($lscale$) and same marginal variance ($\alpha$) as the data-generative function ($True function$).}
  \label{fig7_varing_lscale}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
\caption{The X-axes represent the true data-generative functional lengthscales of every dataset scenario. The Y-axes represent the lengthscales estimates of GP and HSGP models. The use of different $m$ in the HSGP models is represented in separated plots. The dashed black line represents the recommended minimum lengthscales that can be closely approximate by the HSGP model in every case according to Figure \ref{fig3_lscale_vs_J_vs_c}.}
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.45]{fig9_MSE_varing_lscale.pdf}
\caption{RMSE of the HSGP models for the different datasets with different wiggly effects (lengthscales) and number of basis functions $m$.}
  \label{fig9_MSE_varing_lscale}
\end{figure}

We can appreciate an agreement between Figures \ref{fig8_Tlscale_vs_Elscale} and \ref{fig9_MSE_varing_lscale} except for small lengthscales and long lengthscales. We observe small RMSE for large lengthscales, however we appreciate mismatch between large lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}. This is because big changes on large lengthscales, which might be clearly appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, might result in similar posterior and therefore a small RMSE. On the other side, we observe high RMSE for small lengthscales, however we appreciate match between small lengthscale estimates of GP and HSGP models in Figure \ref{fig8_Tlscale_vs_Elscale}. This is because very small changes on small lengthscales, which might not be appreciated in Figure \ref{fig8_Tlscale_vs_Elscale}, can have big effects on the posteriors and therefore high RMSE.

The dashed black line in Figure \ref{fig8_Tlscale_vs_Elscale} represents the minimum lengthscales that can be closely approximate in every case, according to Figure \ref{fig3_lscale_vs_J_vs_c}. It can be appreciated a complete agreement between those recommended minimum lengthscales that can be properly fitted in every case following Figure \ref{fig3_lscale_vs_J_vs_c} and the real approximate accuracy measured by the RMSE in Figure \ref{fig9_MSE_varing_lscale}.
